---
title: "BeginnR Fall 2021: Activity 4 (due 11/8/2021)"
output: html_document
---

Instructions: Please submit the completed .Rmd file via the Google Form linked in Slack. Rename the file as LASTNAME_FIRSTNAME.Rmd, replacing LASTNAME and FIRSTNAME with your last name and first name, respectively.

0. Recap

You're going to want to run this chunk :)
```{r}
library(tidyverse)
library(nycflights13)
```

1. In the `diamonds` tibble, explain whether each of the following variables is categorical or quantitative. Give a short justification for each.
- `carat`: quantified through medians and means, quantitative
- `cut`: categorized from Fair to Ideal, categorical
- `color`: categorized from D to J, categoircal 
- `clarity`: categorized from SI1 to VVS1, categorical
- `table`: quantified through medians and means, quantitative
- `price`: quantified through medians and means, quantitative

```{r}
summary(diamonds)
```

2. Make an appropriate plot showing the **proportion** of diamonds of each color. Hint: there are many ways to do this. You can use the Internet if you get stuck, but it should be doable with methods covered in BeginnR so far.

```{r}
diamonds %>%
  ggplot(mapping = aes(x=color))+
  geom_bar()
```


3. Make an appropriate plot showing the distribution of price, broken down by the `cut`. Your partner doesn't care about diamonds of Fair or Good cut, so only show the distribution for Very Good, Premium, and Ideal cuts.

```{r}
diamonds %>%
  ggplot(mapping = aes(x=cut, y=price))+
  geom_boxplot()
```


4. In Demo 4, it is mentioned that `geom_hex()` is likely better than `geom_point()` when there are a lot of points. Why is this the case?

The density of the points would obscure the countability of the points, making it difficult to see how many points there actually are.

5. When would you want to use `geom_jitter()` instead of `geom_point()`? If you were to plot `x` versus `y` in the `diamonds` tibble, would you use `geom_jitter()`, `geom_point()`, or `geom_hex()`?

geom_jitter is a useful way of handling overplotting caused by discreteness in smaller datasets.

```{r}
diamonds%>%
  ggplot(mapping=aes(x,y))+
  geom_hex()
```
I would use geom_hex because it is difficult to see through the density of the points with the other methods of graphing.

6. You're looking to spend between $3200 and $3300 (inclusive) on a diamond. Make a scatterplot of the `y` vs. the `x` of these diamonds. Make the points on the plot filled diamonds, with the fill color based on the color of each diamond, and make the size (area) of the diamond-shaped points proportional to each diamond's `carat` value. Hint: https://ggplot2.tidyverse.org/articles/ggplot2-specs.html and/or other online resources may be helpful.

```{r}
diamonds%>%
  filter(price >= 3200 & price <=3300)
  ggplot(mapping=aes(x=x, y=y))+
  geom_point(fill = color, size = carat)
```


7. Add a line of best fit to the plot in question 6 using `geom_smooth()`. Color it red, and do NOT display the grey standard error region.

```{r}
diamonds%>%
  filter(price >= 3200 & price <=3300)
  ggplot(mapping=aes(x=x, y=y))+
  geom_point()+
  geom_smooth()
  
```


8. Pick any of the plots in the previous questions, and create an appropriate title and axis labels. Make the title centered.

```{r}
diamonds %>%
  ggplot(mapping = aes(x=color))+
  geom_bar()+
  labs(
    title = "Proportion of Diamonds of Each Color",
    X = "Color",
    Y = "Proportion of Diamonds",+
  theme(plot.title = element_text(hjust = 0.5))
  )
```



