---
title: "BeginnR Fall 2021: Activity 3 (due 11/1/2021)"
output: html_document
---

Instructions: Please submit the completed .Rmd file via the Google Form linked in Slack. Rename the file as LASTNAME_FIRSTNAME.Rmd, replacing LASTNAME and FIRSTNAME with your last name and first name, respectively.

0. Recap

You're going to want to run this chunk :)
```{r}
library(tidyverse)
library(nycflights13)
```


1. What is the difference between the dplyr functions `mutate()` and `transmute()`?

```{r}
?transmute
```
mutate() adds new variables and preserves existing ones; transmute() adds new variables and drops existing ones.

The next few questions will use the `nycflights13::flights` dataset

2. Write code to select all flights that:
(a) Had an arrival delay of two or more hours
(b) Flew to the Bay Area (SFO, SJC, OAK, PAO, STS):
(c) Arrived more than two hours late, but didn’t leave late:
(d) Departed between midnight and 6am (inclusive):
(e) Were delayed by at least an hour, but made up over 30 minutes in flight:
(f) Were operated by United, American, or Delta. Hint: Check the help menu! 

```{r}
flights
```
a)

```{r}
flights%>%
  filter(arr_delay>=2)
```
b)
```{r}
flights%>%
  filter(dest == SFO|SJC|OAK|PAO|STS)
```
c)
```{r}
flights %>%
  filter(arr_delay>=2 & dep_delay<=0)
```
d)
```{r}
flights %>%
  filter(dep_time)
```
e)
```{r}
flights %>%
  mutate(gain = sched_arr_time - arr_time)
  filter(arr_delay>=1 & gain>=30)
```
f)
```{r}
flights %>%
  filter(carrier)%>%
  airlines(United|American|Delta)
```


3. Calculate the average speed each plane was traveling during its journey. Call this variable `speed` and make it in mph.

```{r}
flights%>%
  mutate(
    speed = distance/(arr_time - dep_time)
  )
```


4. Create a histogram showing the distribution of the speed of flights departing JFK airport in April.

```{r}
flights%>%
  filter(month == 4 & origin == "JFK")%>%
  ggplot(mapping=aes(x = speed))
  geom_histogram()
```


5. Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Overwrite them so that they give the number of minutes since midnight. Hint: recall the integer division operator `%/%`. You can also use `floor()`.

6. Modify the code below to compute the departure delay manually using the variables `dep_time` and `sched_dep_time`, and save the result in a new variable `dep_delay_2.` How does `dep_delay_2` compare with `dep_delay`? Is `dep_delay` consistent with `dep_time` and `sched_dep_time`? 


```{r}
new_flights <- 
  flights %>%  
  mutate(dep_delay_2 = sched_dep_time)
```


Use this code to look at the two columns side by side. 
```{r}
new_flights %>% 
  select(dep_delay, dep_delay_2)
```

Use this code chunk to see if yours is identical or not.
```{r}
identical(new_flights$dep_delay, new_flights$dep_delay_2)
```

7. Replace the `NA`s below to define variable called `flight_distances` and `flight_carriers`. Each of these variables should be a vector containing the data from the appropriate column of the `flights` tibble.

```{r}
flight_distances <- NA
flight_carriers <- NA
```


8. Explain, in words, the data in `flight_subset` as defined below.

```{r}
flight_subset <- flight_distances[!(flight_distances < 500) & (flight_carriers %in% c("UA", "AA", "DL"))]
flight_subset
```
Hint: you might want to try examples with your own, probably smaller vectors, or use Google :)


9. Write code that uses dplyr verbs and the pipe operator to obtain the same data as in `flight_subset`, starting from the `flights` tibble. Check your answer using `identical()`. We've gotten you started below.

```{r}
flight_subset_2 <- flights
```


__Congratulations on making it through 3 weeks! You're well on your way to becoming a data wrangling master!__

# 3) Supplemental Reading

https://r4ds.had.co.nz/transform.html

https://dcl-2019-04.github.io/curriculum/manip-basics.html
